---
title: "CM2605-Simulation and Modelling Techniques"
author: "Lovindya Hangilipola"
date: "2024-03-17"
output:
  
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

-   **Mean Arrival Rate:** 1 customer per minute
-   **Mean Service Rate:** 1/3 customers per minute (service time = 3 minutes)
-   **Simulation Length:** 24 hours (converted to minutes: 60 \* 24)

## Simulation Process

The simulation iterates through various numbers of Customer Service Representatives (CSRs) (1 to 20). For each iteration:

1.  **Set the number of CSRs:** The current staffing level is defined.
2.  **Initialize the simulation state:**
    -   Arrival times are generated using Poisson distribution with a mean arrival rate.
    -   Service times are generated using exponential distribution with a mean service rate.
    -   Initial queue and current time are set to empty and 0, respectively.
3.  **Simulate the call center:**
    -   The simulation iterates through each minute of the simulation length.
        -   **Check for arrivals:** If the current time coincides with an arrival time, a customer enters the queue.
        -   **Check for departures:** If a CSR is available (queue is not empty) and the first customer's service time has elapsed, the customer departs, and their waiting time is recorded.
        -   **Update the current time:** The simulation moves to the next minute.
4.  **Calculate the average waiting time:** The average waiting time for all customers served during this staffing level is calculated.

## Results and Analysis

The simulation generates a relationship between the number of CSRs and the average waiting time for customers. This relationship can be visualized using a plot.

### Set the simulation parameters

```{r}
lambda <- 1  # Mean arrival rate (customers per minute)
mu <- 1/3  # Mean service rate (customers per minute)
sim_length <- 60 * 24  # Simulation length (minutes)
```

### Initialize variables

```{r}
wait_times <- c()
num_csrs <- c()
avg_wait_time <- c()
```

### Simulate different number of CSRs

```{r}
# Simulate different number of CSRs
for (i in 1:20) {
  # Set the number of CSRs
  num_csrs[i] <- i

  # Initialize the simulation state
  arrivals <- rpois(sim_length, lambda)
  service_times <- rexp(sim_length, 1/mu)

  # Simulate the call center
  current_time <- 0
  queue <- c()

  for (j in 1:sim_length) {
    # Check for arrivals
    if (current_time >= arrivals[j]) {
      queue <- c(queue, current_time)
    }

    # Check for departures
    if (length(queue) > 0) {
      if (current_time >= queue[1] + service_times[j]) {
        wait_times <- c(wait_times, current_time - queue[1])
        queue <- queue[-1]
      }
    }

    # Update the current time
    current_time <- current_time + 1
  }

  # Calculate the average waiting time
  avg_wait_time[i] <- mean(wait_times)
}


```

### Plotting the results

```{r}
# Plot the results
plot(num_csrs, avg_wait_time, type = 'b', xlab = 'Number of CSRs', ylab = 'Average Waiting Time (minutes)')

```

## Q2

This investigates the power dissipated in a resistor within an electrical circuit.

## Simulation Setup

We assume the voltage (V) across the resistor follows a normal distribution with a mean of 12V and a standard deviation of 2V. Similarly, the resistance (R) follows a normal distribution with a mean of 8 and a standard deviation of 1 .

### Generating Power Values

```{r}
# Generate 7000 values for voltage and resistance
V <- rnorm(7000, 12, 2)
R <- rnorm(7000, 8, 1)

# Calculate power using the formula P = V^2 / R
P <- V^2 / R
```

### Plotting the result

```{r}
# Create a scatter plot
plot(V, P, xlab = "Voltage (V)", ylab = "Power (W)")
title("Scatter Plot of Power vs Voltage")
```

### Find the approximate mean and variance of power (P)

```{r}
#3. Find the approximate mean and variance of power (P).
print("Mean of the power") 
mean(P) 
print("Variuance of the power") 
var(P)
```

### Estimate probability of power exceeding 20W

```{r}
# Estimate probability of power exceeding 20W
probability_high_power <- mean(P > 20)
cat("Probability of power > 20W:", probability_high_power, "\n")
```

# Q3

### 1. Generate N = 500 samples of size n = 50 from a Uniform [−5,5] distribution.

```{r}
N <- 500
n <- 50
samples <- matrix(runif(N * n, -5, 5), nrow = N, ncol = n)
```

### 2. For each of the N = 500 samples, calculate the sample mean, so that you now have a vector of 500 sample means.

```{r}
sample_means <- colMeans(samples)
```

### 3. Plot a histogram of these 500 sample means. Does it look normally distributed and centred at 0?

```{r}
hist(sample_means, main = "Histogram of Sample Means (n = 50)")
```

#### 4. Turn this simulation into a function that takes arguments N the number of simulated samples to make and n the sample size of each simulated sample.

```{r}
simulate_sample_means <- function(N, n) {
  samples <- matrix(runif(N * n, -5, 5), nrow = N, ncol = n)
  sample_means <- colMeans(samples)
  hist(sample_means, main = paste("Histogram of Sample Means (n =", n, ")"))
}
```

#### 5. Run this function for n = 10,15,30,50. What do you notice about the histogram of the sample means (the sampling distribution of the sample mean) as the sample size increases

```{r}
{simulate_sample_means(500, 10)}
```

```{r}
simulate_sample_means(500, 15)
```

```{r}
simulate_sample_means(500, 30)
```

```{r}
simulate_sample_means(500, 50)
```

As the sample size increases, the histogram of the sample means becomes more bell-shaped and centred at 0. This is consistent with the Central Limit Theorem, which states that the sampling distribution of the sample mean will be approximately normal for large sample sizes, regardless of the distribution of the population from which the samples are drawn.

## Q4

```{r}
# returns the random walk path values as a vector 
# (random walk always starts at 0)
# p: probability of increasing by 1
# stop if path value hits either `lower` or `upper`
run <- function(p,lower, upper) {
    values <- c(0)
    current <- 0
    while (current > lower & current < upper) {
        current <- current + ifelse(runif(1) < 1/2, 1, -1)
        values <- c(values, current)
    }
    values
}
```

5)  

```{r}
N <- 5  # no. of walks to simulate
p <- 0.5
lower <- -100
upper <- 100
# simulate paths
set.seed(1055)
vlist <- replicate(N, run(p, lower, upper))

# get length of longest path
max_length <- max(sapply(vlist, length))
# make plot
par(mar = rep(0, 4))  # no margins
plot(c(1, max_length), c(lower, upper), type = "n")
for (i in 1:N) {
    lines(1:length(vlist[[i]]), vlist[[i]])
}
abline(h = 0, lty = "dashed")
abline(h = lower, lwd = 2)
abline(h = upper, lwd = 2)
```

**Ans : As n tends to infinity, the random variable Sₙ will tend to infinity with probability 1. This is because the expected value of Xₙ is 0, so the expected value of Sₙ is also 0. However, the variance of Sₙ will also tend to infinity as n tends to infinity, so Sₙ will become increasingly dispersed as n gets larger**
